\huge{Meta-analytic evidence that sexual selection improves population fitness}

### 2019-03-27 Response_to_editorial_requests.Rmd

---
title: 'Meta-analytic evidence that sexual selection improves population fitness'
author: 
- Justin G. Cally, Devi Stuart-Fox and Luke Holman
output:
  pdf_document: default
subtitle: Response to Editorial Requests and Reviewer
bibliography: Manuscripts/Bibliography/Bibliography.bib
csl: Manuscripts/Bibliography/NComms.csl
header-includes: 
- \usepackage{hyperref}\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, urlcolor=blue} 
---

@@

Dear Dr. Jones and colleagues,

We are delighted to provide a revised manuscript for publication in \textit{Nature Communications}. Below we provide a point-by-point response to the editorial requests as well as a response to comments by Prof. Jacek Radwan.

## Editorial Requests

\begin{enumerate}
  \item \textit{Please supply the main manuscript file in Microsoft Word or LaTeX format. Please supply a Supplementary Information file in PDF format.} \\ 
  The main manuscript is provided in LaTeX format (.tex) and the Supplementary Information is now in pdf format. However, we retain the HTML formatting for the Supplementary Information hosted on \href{https://justincally.github.io/SexualSelection/}{Github}.

  \item \textit{Please edit the title so that it is 15 words or fewer and does not include punctuation} \\
  We have changed our title to: Meta-analytic evidence that sexual selection improves population fitness.

@@

  \item \textit{The final paragraph of the Introduction should summarise the major results and conclusions of this manuscript, in the present tense.} \\
  We have now replaced our aims in the last paragraph of the last paragraph of the introduction with a summary of the major results and colnclusions of the study.

  \item \textit{Please shorten all subheadings in the Results section to fewer than 60 characters including spaces.} \\
  Sexual selection is associated with higher mean values for most fitness components \textbf{to} Sexual selection improves average fitness for most traits. \\
  Sexual selection reduces phenotypic variance, for female traits in stressful environments \textbf{to} Effects of sexual selection on phenotypic variance

  \item \textit{Please do not use italics or bold font to convey emphasis (in both the main text and the display items).} \\
  We only use italics for phrases or abbreviations in Latin (\textit{e.g., i.e., Prima facie}) and not for emphasis.

@@

  \item \textit{Please provide a full Methods section in the main manuscript file. Please note that there are no word limits to the Methods section. The Methods section must include subheadings of fewer than 60 characters including spaces. There is no word limit for the Methods section, so we recommend against splitting the methods description between the main text and Supplementary Information.} \\
  All headings are under 60 characters in length, we have some additional Supplementary Methods in the Supplementary Information but it is an appendage to the Supplementary Tables 1-3 and search terms (which would not nicely format in the main text).

  \item \textit{Please rename the Methods section as "Methods".} \\
  Done

  \item \textit{All Nature Communications manuscripts must include a section titled "Data Availability" as a separate section after the Methods section and before the References.} \\
  We have now ammended this section of the manuscript by seperating data availability and code availability as well as providing the relevent links to the data sources hosted on Github.

  \item \textit{We strongly encourage authors to deposit all new data associated with the paper in a persistent repository where they can be freely and enduringly accessed.} \\
  We have chosen to host our data on a Github repository alongside all our code and a neatly formatted HTML with supplementary information in an effort to maximise transparency of our analysis as well as allowing for the data and code to be easily reproduced and built upon by future studies.

@@

  \item \textit{If you opted into the journal hosting details of a preprint version of your manuscript via a link on our dedicated website (https://nature-research-under-consideration.nature.com), it will remain on this site while you are revising your manuscript, as we consider the file to remain active. Should you wish to remove these details, please email naturecommunications@nature.com indicating your manuscript number and the link on our website that was previously sent to you. Please see our pre-publicity policy at http://www.nature.com/authors/policies/confidentiality.html For more information, please refer to our FAQ page at https://nature-research-under-consideration.nature.com/posts/19641-frequently-asked-questions} \\
  Not applicable to our manuscript.

\end{enumerate}

## Response to Prof. Jacek Radwan

\begin{enumerate}
  \item \textit{The authors made a good effort to address my concerns. In particular, they now report a separate analysis addressing the effect of sexual selection manipulation on traits with direct effect on population fitness, which is the most relevant analysis for the question being asked. This effect has proved ambiguous (significant with REML, but with 95\%Cis overlapping zero with Bayesian analysis), which, coupled with some evidence for publication bias (lines 147-151) , makes the main conclusion of positive effect of selection on population fitness rather tentative. This interpretational uncertainty, however, is not properly reflected in the revised manuscript, eg. in the abstract (l. 12-13 ) and discussion (172-175; 249-251). Furthermore, recommendations for conservation biology seem premature in this context.}\\
  We have modified the title and abstract of the paper, and made it clear which conclusions are supported unambiguously and which are more tentative. For example…

\begin{itemize}

\item The title is now \textit{Meta-analytic evidence that sexual selection improves population fitness}.

\item In the abstract we include: \textbf{We find evidence} that sexual selection on males tends to elevate the mean and reduce the variance for many fitness traits.

\item In the abstract we include: However, sexual selection had comparatively less positive effects on traits directly related to fitness than those with indirect or ambiguous relationships.

\item We have modified the 2nd, 3rd and 4th paragraphs of the discussion and take a more tentative approach in drawing conclusions (see following comments).

\end{itemize}

@@

  \item \textit{Sorry for not being clear with my point on random slopes, I meant the first option the authors identified, but (as I already indicated in my original review) I appreciate that dataset at hand may make this difficult. I’m OK with the way this comment was taken into account.}\\
  We are glad that the reviewer found the modifications we made satisfactory.

  \item \textit{I appreciate the added text explaining that stronger response in females is counterintuitive (l. 195-222). However, it is still not clear what the authors mean to say about the nature of selection on females. I can see two options (which the authors also identify in their response to my question) (i) stronger correlated response in females to selection on males, compared to response in males themselves and (ii) direct effect of sexual selection manipulation on female fitness, e.g. due to increased male harassment. These two mechanisms should be clearly separated and discussed, they can have different consequences for interpretations of the results, but also on recommendations about allowing for “good genes” sexual selection in endangered species (eg. line 243).}\\
  We agree that this section of the discussion can be clearer and have modified the two paragraphs in question.

The section in questions rests upon the fundamental premise that the response to selection depends on A) genetic variance, and B) the strength of selection. This fact is captured in the Breeder’s equation, $R = h^2s$, where $R$ is the response to selection, $h^2$ is the narrow-sense heritability of the trait, and s is the selection differential. Because females responded more strongly than males did (bigger $R$ for females), our argument is that $h^2$ and/or $s$ must be bigger for females. We then list a number of mechanisms that might make $h^2$ or $s$ bigger for females.

\end{enumerate}
Here are the revised paragraphs:

The results of the meta-analysis support predictions that sexual selection on males can improve population fitness and accelerate adaptation [@RN94; @RN173; @RN54; @RN141; @RN230; @RN198; @RN58]. One possible mechanism is that male mating success is positively genetically correlated with traits that contribute to population fitness, allowing females to benefit from a genome that has been purged of deleterious alleles through competition between males [@RN94; @RN198; @RN95]. A second (non-exclusive) mechanism is that experimental manipulation of sexual selection on males might directly alter the selective pressures acting on females, causing female traits to evolve. For example, removing sexual selection via enforced monogamy probably alters selection on females, because it alters the frequency of interactions with males (as well as the evolved genotype of those males). What is less clear is why the manipulation of sexual selection had a larger effect on female trait means and variances as opposed to males -- this result is arguably the opposite of what one would expect, since it is males that experience stronger sexual selection. Below we discuss possible explanations for this result, in light of the core principle that the extent of adaptation depends on additive genetic (co)variance and the strength of selection [@RN317; @RN318].

Firstly, it is possible that female traits show more additive genetic variation than male traits, causing female traits to respond more strongly to a change in selection. This hypothesis is plausible because males frequently do experience stronger selection than females [@Singh2018], and sustained strong selection reduces heritability. A systematic review found no overall difference in mean heritability between male and female traits [@RN342], but did record numerous instances in which trait heritability was higher for females than males [@RN342]. The sex chromosomes provide another reason for sex-specific heritability. Males are heterogametic in most of the species in our sample (*i.e.* the species with XY or XO sex determination), which can reduce father-to-son heritability relative to mother-to-daughter, since sons do not inherit the larger sex chromosome from their fathers [@RN340; @RN262; @RN147; @RN263], potentially slowing the adaptation of male traits [@RN263; @RN259].

Secondly, selection on males might be weaker than selection on females, resulting in slower adaptation following the experimental manipulation of sexual selection. This explanation may initially seem implausible, because net selection on males is often stronger than on females [@Singh2018], due in part to the elevated importance of sexual selection in males as opposed to females [@RN304; @RN303; @RN337; @RN338]. However, an oft-overlooked aspect is that selection might frequently be softer on males and harder on females [@RN319], because the local competitive environment is usually more important for males than it is for females. For instance, a mediocre male genotype can have high fitness provided it outcompetes its local rivals, while low-fitness female genotypes might produce few offspring even when competing with other low-fitness females. Therefore, improvements in genetic quality might have stronger diminishing returns in males, possibly contributing to our finding that the genetic consequences of sexual selection lead to greater fitness benefits for females. Though this argument is speculative, we note that many experimental evolution designs exaggerate the sex difference in the softness of selection, relative to expectations for large, natural populations [@RN322; @RN277; @RN112]. For example, many studies [@RN110; @RN326; @RN300; @RN111; @RN227] have evolved insects in small sub-populations, each containing one female and multiple males, whose progeny are then mixed and randomly sampled to create the next generation; this design ensures that successful males simply needed to outcompete their rival(s) in the same sub-population (soft selection), while each female's reproductive output is measured against the entire female population (hard selection).

@@

\begin{enumerate}
\setcounter{enumi}{3}
  \item \textit{l. 21: I wonder if defining population fitness as an average (presumably in both sexes) makes sense? As the authors state later in abstract, what really matters is fitness of demographically-limiting sex (and the traits which affect the demography).}\\
  We have not made a change here, for two reasons.

Firstly, it usually makes no difference if one thinks of population fitness as an average across all the females, or an average across both sexes. This is because the average fitness of males is always equal to the average fitness of females, assuming the population has obligate sexual reproduction and a 50:50 primary sex ratio (as most species in our meta-analysis do), because every offspring has one genetic mother and one genetic father. For example, if the average fitness across all the females is 1.1, the average fitness of all the males must be 1.1 as well, so one gets the same answer when averaging over females only, or the whole population.

Secondly, ‘population fitness’ has a prominent technical definition that has been used by eminent figures in population geneticists (e.g. Wright, Fisher and Kimura), i.e. the average across the whole population. However, many of these classic models implicitly ignored males (e.g. by assuming isogamous sexual reproduction) for reasons of mathematical tractability. Later ecological and evolutionary models have incorporated the sexes explicitly, and unsurprisingly found that adaptation in males and females has different effects on the growth rate of the population (confusingly, the growth rate of the population is also called ‘population fitness’ by many ecologists). We believe that a clear, inclusive, and universally satisfying definition of population fitness is currently missing, but it is beyond the scope of this paper – instead, we use the standard definition.

  \item \textit{l. 162: low effect on sexual selection manipulation on traits like viability or female fecundity, compared to “indirect effect traits” which including male mating success, align with findings of meta-analysis of Prokop et al. 2012 – it would be appropriate to mention this consistency here.}\\
Thank you – we have now added this reference to the sentence in question.

  \item \textit{l. 165: but see recent metaanalysis by Kelly et al. (2018) showing no sex effect on immunity}\\
This study is interesting, but ultimately we don’t think it is sufficiently relevant to cite here. Our results found evidence that the experimental manipulation of sexual selection causes immunocompetence to evolve, while Kelly et al. (2018) examined whether males and females are differently defended against pathogens.

  \item \textit{I hope supplementary figures will be linked to the main text, as it is somewhat difficult to find them by the number in the supplement deposited in a repository.}\\
In the revision, we have provided a more traditional PDF version of the supplementary material, which is a simple list of all the supplementary figures and tables (this was also added to conform to Nature Communications style). For the submitted version of the manuscript we have hyperlinked the Supplementary Figure/Table to the relevant section of the Github page. One can also find a particular figure in the HTML version using Ctrl + F to search for the figure by name.

### 2019-03-27 Manuscripts/Manuscript_MetaAnalysis.Rmd
minor changes to manuscript and response to editor

---
title: '\huge{Meta-analytic evidence that sexual selection improves population fitness}'
author: 
- Justin G. Cally\textsuperscript{§*}, Devi Stuart-Fox\textsuperscript{§} and Luke Holman\textsuperscript{§}
- 
- '\textsuperscript{*}\textit{justin.g.cally@gmail.com}'
output:
  pdf_document: default
  html_document: default
  word_document:
      pandoc_args: ["--smart"]
bibliography: Bibliography/Bibliography.bib
subparagraph: yes
urlcolor: blue
csl: Bibliography/NComms.csl
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
  \rhead[]{Sexual selection improves fitness} \usepackage{titlefoot}
  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\onehalfspacing
    \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{} \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{} \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} 
    \usepackage{hyperref}\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, urlcolor=blue} 
    \usepackage{color,soul}
    \raggedbottom    
abstract: Sexual selection has manifold ecological and evolutionary consequences, making its net effect on population fitness difficult to predict from first principles. Instead, one can empirically determine whether sexual selection increases or decreases population fitness by  manipulating the opportunity for sexual selection in replicate populations, allow evolution to proceed, and then compare the evolved average fitness between the treatment groups. Here, we synthesise 459 effect sizes from 65 pertinent experimental evolution studies using meta-analysis. We find that sexual selection on males tends to elevate the mean and reduce the variance for many fitness traits, especially in females evolving under stressful conditions. We conclude that the beneficial population-level consequences of sexual selection typically outweigh the harmful ones and that the effects of sexual selection on total selection, genetic variance, and/or evolvability can differ markedly between sexes and environments. We discuss the implications of these results for conservation and evolutionary biology. \vspace{5mm} \par\noindent \textbf{Keywords:} Adaptation, Gender load, Monogamy, Mutation load, Sexual conflict.
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}

@@

Our meta-analysis revealed that populations evolving under sexual selection often have higher values for multiple fitness traits, relative to populations where sexual selection on males was experimentally removed or weakened. Sexual selection had beneficial effects for traits that were ambiguously or indirectly related to fitness, but had smaller, variable effects on direct measures of fitness such as extinction rate, reproductive success (defined as the number of offspring produced) and the proportion of viable offspring. Fitness traits related to immunocompetence were an exception: sexual selection typically resulted in weaker immunity. This result is interesting in light of the hypothesised trade-off between sexually-selected phenotypes and immunity, for example, due to immunosuppressive effects of sex hormones [@RN323; @RN324]. Furthermore, the overall benefit of sexual selection was greater for females than males, and this sex difference was magnified in stressful environments. Consistent with stronger selection on female fitness under stress, female and mixed-sex samples showed reduced phenotypic variance when sexual selection was applied under stressful as opposed to benign conditions. These results suggest that sexual selection may contribute to population persistence under stressful conditions, such as fluctuating environmental change [@RN173] or spatial variability [@RN189], particularly since female reproductive output is often a limiting factor in population growth [@RN188].

The results of the meta-analysis support predictions that sexual selection on males can improve population fitness and accelerate adaptation [@RN94; @RN173; @RN54; @RN141; @RN230; @RN198; @RN58]. Specifically, our results support arguments that positive genetic correlations between male mating success and female non-sexual traits are common, allowing females to benefit from a genome that has been purged of deleterious alleles through sexual selection on males [@RN94; @RN198; @RN95]. What is less clear is why sexual selection had a larger effect on female trait means and variances as opposed to males (the sex in which sexual selection is typically strongest). Below we discuss possible explanations for this result, in light of the core principle that the extent of adaptation depends on additive genetic variance, the structure of genetic covariance, and the strength of selection [@RN317; @RN318]. However, these explanations are speculative and _post hoc_, as the result is arguably opposite to what one would predict, (*i.e.* that manipulating sexual selection on males should affect the phenotype of males more than females).

Firstly, the response of male fitness to sexual selection may be constrained in species where the heritability of male fitness is influenced by the sex determination system. Males are heterogametic in many of the species in our sample (*e.g.* due to XY or XO sex determination); this can reduce father-to-son heritability relative to mother-to-daughter, since sons never inherit the larger sex chromosome from their fathers [@RN340; @RN262; @RN147; @RN263], which might slow adaptation in male relative to female traits [@RN263; @RN259]. However, X-linkage is also likely to inflate the contribution of X-linked genes to fitness variance in the heterogametic sex. Under this view, X-linkage facilitates adaptation by increasing the efficacy of selection against recessive deleterious alleles and selection for recessive beneficial alleles [@RN341; @RN339]. Perhaps as a result of these conflicting predictions, there seems to be no clear consensus regarding sex-based differences in heritability. For instance, a systematic review found no difference in mean heritability between male and female traits, although there was a male-biased skew in heritability of reproductive traits [@RN342]. Conversely, a recent association study in humans found a number of traits showing higher heritability in females than males [@RN325].

Secondly, selection on males might be weaker than selection on females, resulting in slower adaptation following the experimental manipulation of sexual selection. This explanation may initially seem implausible, because net selection on males is generally assumed to be stronger than on females, due to (actual or hypothesised) sex differences in sexual selection and inter-individual variance in fitness [@RN304; @RN303; @RN337; @RN338]. However, an oft-overlooked aspect is that selection might frequently be softer on males and harder on females [@RN319], because the local competitive environment is usually more important for males than it is for females. For instance, a mediocre male genotype can have high fitness provided it outcompetes its local rivals, while low-fitness female genotypes are likely to produce few offspring even when competing with other low-fitness females. Therefore, improvements in genetic quality should have stronger diminishing returns in males, possibly contributing to our finding that the genetic consequences of sexual selection lead to greater fitness benefits for females. Though this argument is speculative, we note that many experimental evolution designs exaggerate the sex difference in the nature of selection, relative to expectations for large, natural populations [@RN322; @RN277; @RN112]. For example, many studies [@RN110; @RN326; @RN300; @RN111; @RN227] have evolved insects in small sub-populations, each containing one female and multiple males, whose progeny are then mixed and randomly sampled to create the next generation; this design ensures that successful males simply needed to outcompete their rival(s) in the same sub-population (soft selection), while each female's reproductive output is measured against the entire female population (hard selection).

@@

![\textbf{Tests for publication bias in the dataset.} Tests of publication bias are mixed and suggest publication bias may be present. (a) Inspection and statistical tests of the funnel plot reveal large amounts of heterogeneity in the dataset with asymmetry from increased low-powered, large effect studies. (b) No significant correlation exists between journal impact factor and effect size, (c) additionally there is no significant correlation between effect size and year of publication when testing for the time-lag bias. For (b) and (c) point size is proportional to the precision of the effect size (*i.e.* the inverse of its variance). For (a), the dashed red lines represent the 95 % CIs of the grand mean estimate for all effect sizes and the black dotted and dashed lines depict the 95 % and 99.8 % CIs for the dataset. The Grey envelopes in (b) and (c) represent the 95 % CI of the linear regression. Source data are provided at \url{https://github.com/JustinCally/SexualSelection}.\label{publication_bias}](Figures/publicationbias.pdf)

>>>>

---
title: '\huge{Meta-analytic evidence that sexual selection improves population fitness}'
author: 
- Justin G. Cally\textsuperscript{§*}, Devi Stuart-Fox\textsuperscript{§} and Luke Holman\textsuperscript{§}
- '\textsuperscript{§}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia'
- '\textsuperscript{*}\textit{justin.g.cally@gmail.com}'
output:
  pdf_document: default
  html_document: default
  word_document:
      pandoc_args: ["--smart"]
bibliography: Bibliography/Bibliography.bib
subparagraph: yes
urlcolor: blue
csl: Bibliography/NComms.csl
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
  \rhead[]{Sexual selection improves fitness} \usepackage{titlefoot}
  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\onehalfspacing
    \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{} \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{} \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} 
    \usepackage{hyperref}\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, urlcolor=blue} 
    \usepackage{color,soul}
    \raggedbottom    
abstract: Sexual selection has manifold ecological and evolutionary consequences, making its net effect on population fitness difficult to predict from first principles. Instead, one can empirically determine whether sexual selection increases or decreases population fitness by manipulating the opportunity for sexual selection in replicate populations, allow evolution to proceed, and then compare the evolved average fitness between the treatment groups. Here, we synthesise 459 effect sizes from 65 pertinent experimental evolution studies using meta-analysis. We find evidence that sexual selection on males tends to elevate the mean and reduce the variance for many fitness traits, especially in females evolving under stressful conditions. However, sexual selection had comparatively less positive effects on traits directly related to fitness than those with indirect or ambiguous relationships. Ultimately, we conclude that the beneficial population-level consequences of sexual selection typically outweigh the harmful ones and that the effects of sexual selection on total selection, genetic variance, and/or evolvability can differ markedly between sexes and environments. We discuss the implications of these results for conservation and evolutionary biology.
---
\pagestyle{plain}

@@

Our meta-analysis revealed that populations evolving under sexual selection often have higher values for multiple fitness traits, relative to populations where sexual selection on males was experimentally removed or weakened. Sexual selection had beneficial effects for traits that were ambiguously or indirectly related to fitness, but had smaller and variable effects on direct measures of fitness such as extinction rate, reproductive success (defined as the number of offspring produced), and the proportion of viable offspring (similar to the results of another meta-analysis [@Prokop_2012]). Fitness traits related to immunocompetence were an exception: sexual selection typically resulted in weaker immunity. This result is interesting in light of the hypothesised trade-off between sexually-selected phenotypes and immunity, for example, due to immunosuppressive effects of sex hormones [@RN323; @RN324]. Furthermore, the overall benefit of sexual selection was greater for females than males, and this sex difference was magnified in stressful environments. Consistent with stronger selection on female fitness under stress, female and mixed-sex samples showed reduced phenotypic variance when sexual selection was applied under stressful as opposed to benign conditions. These results suggest that sexual selection may contribute to population persistence under stressful conditions, such as fluctuating environmental change [@RN173] or spatial variability [@RN189], particularly since female reproductive output is often a limiting factor in population growth [@RN188].

The results of the meta-analysis support predictions that sexual selection on males can improve population fitness and accelerate adaptation [@RN94; @RN173; @RN54; @RN141; @RN230; @RN198; @RN58]. One possible mechanism is that male mating success is positively genetically correlated with traits that contribute to population fitness, allowing females to benefit from a genome that has been purged of deleterious alleles through competition between males [@RN94; @RN198; @RN95]. A second (non-exclusive) mechanism is that experimental manipulation of sexual selection on males might directly alter the selective pressures acting on females, causing female traits to evolve. For example, removing sexual selection via enforced monogamy probably alters selection on females, because it alters the frequency of interactions with males (as well as the evolved genotype of those males). What is less clear is why the manipulation of sexual selection had a larger effect on female trait means and variances as opposed to males -- this result is arguably the opposite of what one would expect, since it is males that experience stronger sexual selection. Below we discuss possible explanations for this result, in light of the core principle that the extent of adaptation depends on additive genetic (co)variance and the strength of selection [@RN317; @RN318].

Firstly, it is possible that female traits show more additive genetic variation than male traits, causing female traits to respond more strongly to a change in selection. This hypothesis is plausible because males frequently do experience stronger selection than females [@Singh2018], and sustained strong selection reduces heritability. A systematic review found no overall difference in mean heritability between male and female traits [@RN342], but did record numerous instances in which trait heritability was higher for females than males [@RN342]. The sex chromosomes provide another reason for sex-specific heritability. Males are heterogametic in most of the species in our sample (*i.e.* the species with XY or XO sex determination), which can reduce father-to-son heritability relative to mother-to-daughter, since sons do not inherit the larger sex chromosome from their fathers [@RN340; @RN262; @RN147; @RN263], potentially slowing the adaptation of male traits [@RN263; @RN259].

Secondly, selection on males might be weaker than selection on females, resulting in slower adaptation following the experimental manipulation of sexual selection. This explanation may initially seem implausible, because net selection on males is often stronger than on females [@Singh2018], due in part to the elevated importance of sexual selection in males as opposed to females [@RN304; @RN303; @RN337; @RN338]. However, an oft-overlooked aspect is that selection might frequently be softer on males and harder on females [@RN319], because the local competitive environment is usually more important for males than it is for females. For instance, a mediocre male genotype can have high fitness provided it outcompetes its local rivals, while low-fitness female genotypes might produce few offspring even when competing with other low-fitness females. Therefore, improvements in genetic quality might have stronger diminishing returns in males, possibly contributing to our finding that the genetic consequences of sexual selection lead to greater fitness benefits for females. Though this argument is speculative, we note that many experimental evolution designs exaggerate the sex difference in the softness of selection, relative to expectations for large, natural populations [@RN322; @RN277; @RN112]. For example, many studies [@RN110; @RN326; @RN300; @RN111; @RN227] have evolved insects in small sub-populations, each containing one female and multiple males, whose progeny are then mixed and randomly sampled to create the next generation; this design ensures that successful males simply needed to outcompete their rival(s) in the same sub-population (soft selection), while each female's reproductive output is measured against the entire female population (hard selection).

@@

![\textbf{Tests for publication bias in the dataset.} Tests of publication bias are mixed and suggest publication bias may be present. (a) Inspection and statistical tests of the funnel plot reveal large amounts of heterogeneity in the dataset with asymmetry from increased low-powered, large effect studies. (b) No significant correlation exists between journal impact factor and effect size, (c) additionally there is no significant correlation between effect size and year of publication when testing for the time-lag bias. For (b) and (c) point size is proportional to the precision of the effect size (*i.e.* the inverse of its variance). For (a), the dashed red lines represent the 95 % CIs of the grand mean estimate for all effect sizes and the black dotted and dashed lines depict the 95 % and 99.8 % CIs for the dataset. The grey envelopes in (b) and (c) represent the 95 % CI of the linear regression. Source data are provided at \url{https://github.com/JustinCally/SexualSelection}.\label{publication_bias}](Figures/publicationbias.pdf)
